CODECHECK: An Open Science initiative for the independent execution of computations underlying research articles during peer review improves reproducibility

### 2021-02-09 codecheck-paper.tex
consistent formatting

\textbf{Isn't CODECHECK what peer review should be doing already?} \\
On the surface, yes, but peer reviewers are overburdened enough and
asking them to do more work around peer review is not likely to
succeed.  Miyakawa described how when he as an editor simply requested
raw data from authors during submission, this led to almost all of
those papers in which data was requested to be withdrawn or rejected
after peer review \cite{miyakawa_no_2020}.  Such basic checks to show
that sufficient data supports the results requires effort from
editors, yet they only rely on the availability of data files and the
content of the paper.  These availability checks can be enhanced by
having more complex CODECHECKs request the code and then execute it.
This might fall within idealistic expectations of peer review, but is
rare.  Establishing a CODECHECK process is largely an acknowledgement
that peer reviewing practices have been unable to adapt to the
challenges of computational papers.  The concept of a CODECHECK, just
as the concepts of reproducible research and Open Science, may be
transitional by nature. If the activities described here as being part
of a CODECHECK become the norm for scientific reviews and
publications, then the initiative will have succeeded. However, the
rapidly changing technology used in research leads us to believe that
a separate role of codechecker and the expertise it brings will be
beneficial in the long term.

\textbf{Should CODECHECK requirements be more demanding?} \\
CODECHECK by design does not require authors to provide (and sustain)
an eternally functional workflow nor suggests a specific software
stack or practical approach.
Creating something
that anyone can reproduce has been called a fool's
errand (Douglas Blank on Twitter,
\url{https://twitter.com/DougBlank/status/1135904909663068165})
and we tend to agree.  However, the package of data, code, and
documentation collaboratively created by authors and codecheckers is a
snapshot of a working analysis that greatly increases the likelihood
of a successful reproduction and the possibility that a workflow can
be extended by third parties in the future, as long as they have
access to suitable resources and matching skill set.  Concrete
implementations of CODECHECK workflows, especially for specific
disciplines, may reify much more helpful guidelines for authors on how
to create reproducibility packages.  Our author-friendly ``low bar''
should not stay low forever, but cultural change takes time and the
encouragement and guidance that CODECHECK, as part of the widely
accepted peer review concept, can provide may eventually allow the bar
to be raised much higher, e.g.,~with executable research
compendia \cite{nust_opening_2017}, ``Whole Tales''
\cite{brinckman_computing_2018}, or continuous analysis
\cite{beaulieu-jones_reproducibility_2017-1}.  However, considering
that missing artefacts and lack of documentation have repeatedly been
identified as key barriers to reproducibility
(e.g.,~\cite{stagge_assessing_2019,nust_improving_2020}), we would not
underestimate the power of a simple check.  For example, ModelDB
curation policies require that only one figure need be manually reproduced
\cite{mcdougal_reproducibility_2016}, but that has not limited the
usefulness nor success of the platform.

@@

\textbf{How are failures during checks handled?}

We have not yet established a process for if a reproduction fails.  This
may simply reflect our biases in pre-selection of case studies.  In
the case that a journal adopts CODECHECK for all submissions, the open
question remains as what to do if a check fails, after exhausting
efforts between author and codechecker to reproduce the workflow.  A
negative comment in a CODECHECK certificate or a failed check does not
necessarily mean the paper or the science is bad (cf. discussion on
negative comments in \cite{everythinghertz123}).  We doubt that
publicly reporting failures (i.e.,~the code would not run) will
increase overall reproducibility, and we recommend sharing the latest
draft of the certificate only with the authors.  One must be careful
how policies on disclosing failures may \textbf{negatively impact the
  acceptance by authors}
%% SJE: what does this mean?
, and also how volunteer efforts may go unnoticed if no report is
published.  The most appropriate way forwards seems to be to give
private feedback for failures and public acknowledgment and
encouragement for successes.
% eLife's new model (see https://osf.io/9cftx/) will publish reports eventually, e.g.,~after preprint has been accepted elsewhere
Rosenthal et~al. \cite{Rosenthal2016b} discuss such incentives for different actors around the
implementation of reproducibility. We see CODECHECK as one way for
organisations to invest in reproducibility by creating incentives
until reproducible computations become the norm.

@@

\textbf{What does it mean for a figure to be reproducible?}

Automatically detecting if a codechecker's results are ``the same'' as
an author's is more challenging than it might appear.  That is why we
do not require results to be \textit{identical} for a CODECHECK to pass
but simply that the code runs and generates output files that the
author claims. Stochastic simulations mean that often we cannot get
exactly the same results, and that even the same versions of libraries
can generate outputs that differ by operating system
\cite{Gronenschild2012-pp}.  While reproducibility practices can
mitigate some of these problems, e.g., by using a seed, the
flexibility of the human judgement is still needed, rather than
bitwise reproducibility.  The codechecker is free to comment on
visible differences in outputs in their report.

>>>>

\textbf{Isn't CODECHECK what peer review should be doing already?}
On the surface, yes, but peer reviewers are overburdened enough and
asking them to do more work around peer review is not likely to
succeed.  Miyakawa described how when he as an editor simply requested
raw data from authors during submission, this led to almost all of
those papers in which data was requested to be withdrawn or rejected
after peer review \cite{miyakawa_no_2020}.  Such basic checks to show
that sufficient data supports the results requires effort from
editors, yet they only rely on the availability of data files and the
content of the paper.  These availability checks can be enhanced by
having more complex CODECHECKs request the code and then execute it.
This might fall within idealistic expectations of peer review, but is
rare.  Establishing a CODECHECK process is largely an acknowledgement
that peer reviewing practices have been unable to adapt to the
challenges of computational papers.  The concept of a CODECHECK, just
as the concepts of reproducible research and Open Science, may be
transitional by nature. If the activities described here as being part
of a CODECHECK become the norm for scientific reviews and
publications, then the initiative will have succeeded. However, the
rapidly changing technology used in research leads us to believe that
a separate role of codechecker and the expertise it brings will be
beneficial in the long term.

\textbf{Should CODECHECK requirements be more demanding?}
CODECHECK by design does not require authors to provide (and sustain)
an eternally functional workflow nor suggests a specific software
stack or practical approach.
Creating something
that anyone can reproduce has been called a fool's
errand (Douglas Blank on Twitter,
\url{https://twitter.com/DougBlank/status/1135904909663068165})
and we tend to agree.  However, the package of data, code, and
documentation collaboratively created by authors and codecheckers is a
snapshot of a working analysis that greatly increases the likelihood
of a successful reproduction and the possibility that a workflow can
be extended by third parties in the future, as long as they have
access to suitable resources and matching skill set.  Concrete
implementations of CODECHECK workflows, especially for specific
disciplines, may reify much more helpful guidelines for authors on how
to create reproducibility packages.  Our author-friendly ``low bar''
should not stay low forever, but cultural change takes time and the
encouragement and guidance that CODECHECK, as part of the widely
accepted peer review concept, can provide may eventually allow the bar
to be raised much higher, e.g.,~with executable research
compendia \cite{nust_opening_2017}, ``Whole Tales''
\cite{brinckman_computing_2018}, or continuous analysis
\cite{beaulieu-jones_reproducibility_2017-1}.  However, considering
that missing artefacts and lack of documentation have repeatedly been
identified as key barriers to reproducibility
(e.g.,~\cite{stagge_assessing_2019,nust_improving_2020}), we would not
underestimate the power of a simple check.  For example, ModelDB
curation policies require that only one figure need be manually reproduced
\cite{mcdougal_reproducibility_2016}, but that has not limited the
usefulness nor success of the platform.

@@

\textbf{How are failures during checks handled?}
We have not yet established a process for if a reproduction fails.  This
may simply reflect our biases in pre-selection of case studies.  In
the case that a journal adopts CODECHECK for all submissions, the open
question remains as what to do if a check fails, after exhausting
efforts between author and codechecker to reproduce the workflow.  A
negative comment in a CODECHECK certificate or a failed check does not
necessarily mean the paper or the science is bad (cf. discussion on
negative comments in \cite{everythinghertz123}).  We doubt that
publicly reporting failures (i.e.,~the code would not run) will
increase overall reproducibility, and we recommend sharing the latest
draft of the certificate only with the authors.  One must be careful
how policies on disclosing failures may \textbf{negatively impact the
  acceptance by authors}
%% SJE: what does this mean?
, and also how volunteer efforts may go unnoticed if no report is
published.  The most appropriate way forwards seems to be to give
private feedback for failures and public acknowledgment and
encouragement for successes.
% eLife's new model (see https://osf.io/9cftx/) will publish reports eventually, e.g.,~after preprint has been accepted elsewhere
Rosenthal et~al. \cite{Rosenthal2016b} discuss such incentives for different actors around the
implementation of reproducibility. We see CODECHECK as one way for
organisations to invest in reproducibility by creating incentives
until reproducible computations become the norm.

@@

\textbf{What does it mean for a figure to be reproducible?}
Automatically detecting if a codechecker's results are ``the same'' as
an author's is more challenging than it might appear.  That is why we
do not require results to be \textit{identical} for a CODECHECK to pass
but simply that the code runs and generates output files that the
author claims. Stochastic simulations mean that often we cannot get
exactly the same results, and that even the same versions of libraries
can generate outputs that differ by operating system
\cite{Gronenschild2012-pp}.  While reproducibility practices can
mitigate some of these problems, e.g., by using a seed, the
flexibility of the human judgement is still needed, rather than
bitwise reproducibility.  The codechecker is free to comment on
visible differences in outputs in their report.
